---
title: Home health care and readmissions after Myocardial Infarction:Analysis of the
  2012 - 2014 NRD database
output:
  word_document: default
  html_document:
    df_print: paged
---

The *rnotebook* contains code and results for analysis of the NRD data. The data has been initially obtained from the NRD files using SAS. After identifying the MI primary diagnosis admissions and then 30-day readmission, further coding performed in R. 

Inclusion criteria:

1. Adult patients > 17 years, with 1o diagnosis of MI. 30- day readmission tracked and coded for.
2. Removed discharges in Dec for each year to remove those patients who cannot complete 30 days time from discharge.


```{r}
library(tidyverse);library(tidylog);library(haven);library(survey);
library(srvyr);library(skimr);library(Matching);library(MatchIt);
library(cobalt);library(twang);library(MASS);library(car);library(lme4);library(nlme)
```


```{r}

setwd("E:/mi_readmit")
# get the data for further coding. Code used is from the earlier paper worked for only 2014.

# get 2012 first

df12 = read_sas("ami_readmission2012.sas7bdat")

glimpse(df12)

# plan to get the other files for 2012 and create a single file 
# then we can combine all 3 years and do desc statistics prior to 
# propensity matching 

names(df12) = tolower(names(df12))

df12.1 = df12 %>% select(age, aweekend, died, discwt, dispuniform,
        dmonth, dx1:dx15, dxccs3:dxccs10, female, hosp_nrd, 
        key_nrd, nrd_stratum, pay1, pr1:pr10, readmit, year)

glimpse(df12.1) # subset to important variables 

hosp = read_sas("/Volumes/Seagate Backup Plus Drive/NRD2012_3/nrd_2012_hospital_v2.sas7bdat")

names(hosp) = tolower(names(hosp))

glimpse(hosp)

hosp2 = hosp %>% select(hosp_nrd, hosp_bedsize, hosp_ur_teach)

# combine this with the main file

df12.2 = left_join(df12.1, hosp2, by = "hosp_nrd")

# now get the cm codes and then we should be done

sev = read_sas("/Volumes/Seagate Backup Plus Drive/NRD2012_3/nrd_2012_severity_v2.sas7bdat")

names(sev) = tolower(names(sev))

glimpse(sev)

sev2 = sev %>% select(contains("cm"), key_nrd)

glimpse(sev2)

df12.3 = left_join(df12.2, sev2, by = "key_nrd")

glimpse(df12.3)

write_csv(df12.3, "df2012_tojoin.csv")
```

Now do the same for 2013. The 2014 already contains all the *var* and need to remove those not needed.

```{r}
# 2013

df13 = read_sas("ami_readmission2013.sas7bdat")

names(df13) = tolower(names(df13))

glimpse(df13)

df13.1 = df13 %>% select(contains("cm"),
                        age, aweekend, died, discwt, dispuniform,
        dmonth, dx1:dx15, dxccs3:dxccs10, female, hosp_nrd, 
        key_nrd, nrd_stratum, pay1, pr1:pr10, readmit, hosp_ur_teach,
        hosp_bedsize, year)


write_csv(df13.1, "df2013_tojoin.csv")

```

Do the same for 2014

```{r}
# get the main file for 2014

df14 = read_sas("ami_readmission2014.sas7bdat")

names(df14) = tolower(names(df14))

df14.1 = df14 %>% select(age, aweekend, died, discwt, dispuniform,
        dmonth, dx1:dx15, dxccs3:dxccs10, female, hosp_nrd, 
        key_nrd, nrd_stratum, pay1, pr1:pr10, readmit, year)

hosp = read_sas("/Volumes/Seagate Backup Plus Drive/NRD_UH/NRD_2014/hosp_2014.sas7bdat")

names(hosp) = tolower(names(hosp))

glimpse(hosp)

hosp2 = hosp %>% select(hosp_nrd, hosp_bedsize, hosp_ur_teach)

df14.2 = left_join(df14.1, hosp2, by = "hosp_nrd")

sev = read_sas("/Volumes/Seagate Backup Plus Drive/NRD_UH/NRD_2014/sev2014.sas7bdat")

names(sev) = tolower(names(sev))

sev2 = sev %>% select(contains("cm"), key_nrd)

df14.3 = left_join(df14.2, sev2 , by = "key_nrd")

write_csv(df14.3, "df2014_tojoin.csv")
```


Combine all 3 years together.

```{r}

df12.3 = read_csv("df2012_tojoin.csv")

df = rbind(df12.3, df13.1, df14.3) # all combined, still need to remove those that died during the hospital stay

write_csv(df, "combineddf.csv")
```

1. Remove died in hospital
2. Determine discharge status and then get home/hhc groups.
3. Now get the other variables needed.
4. Create tableone and basic survey analysis of cohort
5. Then determine plan for logistic regression model for propensity score

```{r}
# get the combined data 

setwd("H:/mi_readmit/")

df = read_csv("combineddf.csv")

# create survey object so that we can detemine how many removed
# died

library(survey)
options(survey.lonely.psu = "adjust")

# died
df %>% count(died)


df = df %>% drop_na(died) # then remove those that died

# remove died



# remove died == 1 as we want patients who survived admission

dfs = survey::svydesign(data = df, ids = ~hosp_nrd, strata = ~nrd_stratum,
                        nest = T, weights = ~discwt)


svytotal(~factor(died), dfs) # % of patients that died in admission (46915/463124) died during admission
svymean(~died, dfs) # 9.19% died during the hospital stay

# remove those that died

df1 = df %>% filter(died == 0) # keep alive at discharge

df1 %>% count(dispuniform)

# keep only 1 = home and 5 = hhc in the data


df2 = df1 %>% filter(dispuniform %in% c(1,5))

df2$discharge[df2$dispuniform == 1]<- "home"
df2$discharge[df2$dispuniform == 5]<- "hhc"

df2 %>% count(discharge) # home or hhc discharge 

# now plan to do further variables 

# create variables for cabg, pci and thrombolytic therapy

thrombo <- as.character(c("9910")) # thrombolytic therapy

df2$thrombo <- apply(df2[paste0("pr", 1:10)], 1, 
                     function(x) any(x %in% thrombo)) %>% 
  as.numeric()

table(df2$thrombo)
```



```{r}
# pci done during admission

pci = as.character(c('3601', '3602', '3603', '3604', '3606', '3607',
                     '3609', '0066', '0045', '0047', '0048'))


df2$pci <- apply(df2[paste0("pr", 1:10)], 1, 
                     function(x) any(x %in% pci)) %>% 
  as.numeric()

table(df2$pci)
```


```{r}
# primary cabg done during same admission

cabg = as.character(c('3610', '3611', '3612', '3613', '3614', '3615', '3617', '3619'))

df2$cabg = apply(df2[paste0("pr", 1:10)], 1, 
                 function(x) any(x %in% cabg)) %>%
  as.numeric()


df2 %>% count(cabg)
```

main 3 treatment modalities done, now to get the main covariates and comorbidities from icd9 dx codes:

```{r}
# smoking 

smoke = as.character(c('V1582', '3051'))

df2$smoke = apply(df2[paste0("dx", 3:10)], 1, 
                      function(x) any(x %in% smoke)) %>% 
  as.numeric()

df2 %>% count(smoke)
```

```{r}

# dyslipedemia

dys = as.character(c('2720', '2721', '2722', '2723', '2724'))

df2$dys = apply(df2[paste0("dx", 3:10)], 1, 
                function(x) any(x %in% dys)) %>%
  as.numeric()

df2 %>% count(dys)

```
```{r}
# prior mi

priormi = as.character(c('412'))

df2$priormi = apply(df2[paste0("dx", 3:10)], 1, 
                    function(x) any(x %in% priormi)) %>%
  as.numeric()


df2 %>% count(priormi)
```
```{r}
# prior pci

priorpci = as.character(c('V4581'))

df2$priorpci = apply(df2[paste0("dx", 3:10)], 1, 
                     function(x) any(x %in% priorpci)) %>% 
  as.numeric()

df2 %>% count(priorpci)
```
```{r}

# carotid disease 

car = as.character(c('43310'))

df2$car = apply(df2[paste0("dx", 3:10)], 1, 
                function(x) any(x %in% car)) %>% 
  as.numeric()

df2 %>% count(car)



```
```{r}
# prior stroke 

pstroke = as.character(c('V1254', '2380'))

df2$pstroke = apply(df2[paste0("dx", 3:10)], 1, 
                    function(x) any(x %in% pstroke)) %>% 
  as.numeric()

df2 %>% count(pstroke)
```
Understand the variables that we have till now and subset the data to make it more meaningful

```{r}

# for cancer, combine cm_mets and cm_tumor

df2$cancer = with(df2, ifelse((cm_tumor == 1 | cm_mets == 1), 1, 0))

df2 %>% count(cancer)

df2$diabetes = with(df2, ifelse((cm_dm == 1 | cm_dmcx == 1), 1, 0))

df2 %>% count(diabetes)


```

```{r}
glimpse(df2)

# create and clean some variables prior to analysis

table(df2$pay1)

df2$pay1 = as.factor(df2$pay1)

df2$insu[df2$pay1 == "1"]<- "medicare"
df2$insu[df2$pay1 == "2"]<- "medicaid"
df2$insu[df2$pay1 == "3"] <- "private"
df2$insu[df2$pay1 == "4"]<- "others"
df2$insu[df2$pay1 == "5"]<- "others"
df2$insu[df2$pay1 == "6"]<- "others"

table(df2$insu) # primary insurance 


# urban/rural

df2$hosp_loc[df2$hosp_ur_teach == "1"]<- "urban"
df2$hosp_loc[df2$hosp_ur_teach == "0"]<- "urban"
df2$hosp_loc[df2$hosp_ur_teach == "2"]<- "rural"


# finally also create a new variable treatment
# if cabg == 1 then always cabg 
# if cabg == 0 then pci if pci == 1
# lastly thrombolytic therapy == 1


df2$treatment[df2$cabg == 1]<- "cabg"
df2$treatment[df2$cabg == 0 & df2$pci == 1]<- "pci"
df2$treatment[df2$cabg == 0 & df2$pci == 1]<- "pci"
df2$treatment[df2$cabg == 0 & df2$pci == 0 & df2$thrombo == 1]<- "medical"

df2$treatment[is.na(df2$treatment)]<- "medical"

df2 %>% count(treatment)

keep = c("age", "aweekend", "died", "discwt", 
"female", "hosp_nrd", "key_nrd", "nrd_stratum", "pay1", 
"readmit", "hosp_bedsize", "hosp_ur_teach",  "cm_chf", "cm_chrnlung", 
 "diabetes", "cm_htn_c", 
"cm_hypothy", "cm_liver","treatment", 
"cm_obese",  "cm_perivasc", "year",
"cm_renlfail",  "cancer","insu", "hosp_loc", 
"cm_wghtloss", "discharge", "thrombo", "pci", "cabg", "smoke", 
"dys", "priormi", "priorpci", "car", "pstroke")

df3 = df2[,keep] # keep only the important variables 
```


All variables done. Now to create a table one for all the data and then:

1. Unadjusted analysis
2. Propensity matching
3. Adjusted analysis


```{r}
library(tableone)

vars = c("age", "aweekend", "died",
"female", 
"readmit", "hosp_bedsize",   "cm_chf", "cm_chrnlung", 
 "diabetes", "cm_htn_c", 
"cm_hypothy", "cm_liver", "treatment",
"cm_obese",  "cm_perivasc", "year",
"cm_renlfail",  "cancer","insu", "hosp_loc", 
"cm_wghtloss", "discharge", "thrombo", "pci", "cabg", "smoke", 
"dys", "priormi", "priorpci", "car", "pstroke")

factorvars = c( "aweekend", "died",
"female", "treatment",
"readmit", "hosp_bedsize",   "cm_chf", "cm_chrnlung", 
 "diabetes", "cm_htn_c", 
"cm_hypothy", "cm_liver", 
"cm_obese",  "cm_perivasc", "year",
"cm_renlfail",  "cancer","insu", "hosp_loc", 
"cm_wghtloss", "discharge", "thrombo", "pci", "cabg", "smoke", 
"dys", "priormi", "priorpci", "car", "pstroke")

design = survey::svydesign(data = df3, ids = ~hosp_nrd, strata = ~nrd_stratum,
                           nest = T, weights = ~discwt)


t1 = tableone::svyCreateTableOne(vars = vars, factorVars = factorvars, data = design,
                                 strata = "discharge")

table1 = print(t1)

write.csv(table1, "table_one.csv")
```


```{r}

# unadjusted svyglm result

df4$discharge = as.factor(df4$discharge)

df4$discharge = ordered(df4$discharge, levels = c("hhc", "home"))


unadjust = svyglm(readmit ~ hhc, design = sdesign, family = "binomial"(link = "logit"))



# unadjusted risk difference

betaiptw = coef(unadjust)

library(sandwich)

se = sqrt(diag(vcovHC(unadjust, type = "HC1")))

causalrd = (betaiptw[2])

lcl = (betaiptw[2] - 1.96*se[2])

ucl = (betaiptw[2] + 1.96*se[2])

unadjust = c(lcl, causalrd, ucl)

unadjust
```

This is the unadjusted analysis; this demonstrates that patients that have home discharge have much lower rates of readmision compared to patients with 
home health care.

However, we feel that comorbidities influence the readmission rather than actual home or home health care choices. Thus, we feel that we need to determine what is the relationship between home and home health care when we adjust for all the comorbidities. 

treated: home health care
control: home 

We want to see what is the influence of home health care on readmission independent of all associated comorbidities.


```{r create propensity model}

# we will first create a small subset of the data selecting only the variables
# that we need
# then we will perform simple imputation to fill the missing data once we identify those variables that have missing data
# we will then do a svyglm to run the logistic regression model
# using the ATT weights, we will obtain ps and then weights on ps score
# we will calculate the sd before and after weighting and look at balance
# we will then do the svyglm with adjustment
# we will calculate RR using the svyglm and then get robust estimators to obtain adjusted risk ratios after weighting.

# create a subset of data

varstokeep = c("age", "aweekend", "discwt", "female", "hosp_nrd", 
"key_nrd", "nrd_stratum",  "readmit", "hosp_bedsize", 
 "cm_chf", "cm_chrnlung", "diabetes", "cm_htn_c", 
"cm_hypothy", "cm_liver", "treatment", "cm_obese", "cm_perivasc", 
"year", "cm_renlfail", "cancer", "insu", "hosp_loc", "cm_wghtloss", 
"discharge",  "smoke", "dys", "priormi", 
"priorpci", "car", "pstroke")

df4 = df3[,varstokeep]


write_csv(df4, "dftouse.csv")

library(purrr)

miss = purrr::map(df4, ~sum(is.na(.)))

# missing data

miss 

# insurance has only missing data 364/176582 (0.2% missing); imputed by adding to largest group

df4$insu[is.na(df4$insu)]<- "medicare"

df4 %>% count(insu)

write_csv(df4, "dftouse.csv")

# impute insurance and add all to the medicare as that is largest group
```
now create svyglm model to for propensity score
propensity score should be formed using variables that impact the treatment rather than the outcome (i.e. hhc use rather than readmission)

```{r}


df4 = read_csv("dftouse.csv")

df4$hhc[df4$discharge == "hhc"]<- 1
df4$hhc[df4$discharge == "home"]<- 0


df4 = data.frame(df4)

options(survey.lonely.psu = "adjust")

sdesign = survey::svydesign(data = df4, ids = ~hosp_nrd, weights = ~discwt, 
          strata = ~nrd_stratum, nest = T)

psmodel = svyglm(hhc ~ hosp_nrd + age + aweekend + female + cm_chf + cm_chrnlung +  diabetes + cm_htn_c + cm_hypothy + cm_liver + cm_obese + cm_perivasc +
cm_renlfail + cancer + hosp_loc + insu + smoke + year +
dys + priormi + treatment + priorpci + car +
pstroke, design = sdesign, family = "binomial"(link = "logit"))


summary(psmodel)

exp(coef(psmodel))
exp(confint(psmodel))
```

create propensity scores

```{r}

df4$ps = psmodel$fitted
df4$linps = psmodel$linear.predictors

# create the ATT using the ps

df4$wts1 = ifelse(df4$hhc ==1, 1, df4$ps/(1-df4$ps)) # get ATT weights

ggplot(df4, aes(x = ps, y = wts1, color = discharge)) + geom_point() # plot the weights for the control and treat cohort
```


ATT adjusted model with weights


create  table 2 with new weights

```{r}

satt = survey::svydesign(data = df4, ids = ~hosp_nrd, weights = ~discwt + wts1, 
          strata = ~nrd_stratum, nest = T)

vars = c("age", "aweekend", "died",
"female", 
"readmit", "hosp_bedsize",   "cm_chf", "cm_chrnlung", 
 "diabetes", "cm_htn_c", 
"cm_hypothy", "cm_liver", "treatment",
"cm_obese",  "cm_perivasc", "year",
"cm_renlfail",  "cancer","insu", "hosp_loc", 
"cm_wghtloss", "discharge", "thrombo", "pci", "cabg", "smoke", 
"dys", "priormi", "priorpci", "car", "pstroke")

factorvars = c( "aweekend", "died",
"female", "treatment",
"readmit", "hosp_bedsize",   "cm_chf", "cm_chrnlung", 
 "diabetes", "cm_htn_c", 
"cm_hypothy", "cm_liver", 
"cm_obese",  "cm_perivasc", "year",
"cm_renlfail",  "cancer","insu", "hosp_loc", 
"cm_wghtloss", "discharge", "thrombo", "pci", "cabg", "smoke", 
"dys", "priormi", "priorpci", "car", "pstroke")

t2 = tableone::svyCreateTableOne(vars = vars, factorVars = factorvars, data = satt,
                                 strata = "discharge")

table2 = print(t2)

# write.csv(table1, "table_one.csv")

```




```{r}
# with ATT weights

satt = survey::svydesign(data = df4, ids = ~hosp_nrd, weights = ~discwt + wts1, 
          strata = ~nrd_stratum, nest = T)

```





```{r}
# using the twang library for the same 

library(twang)

covlist = c('age' , 'aweekend' , 'female' , 'cm_chf' , 'cm_chrnlung' ,
  'diabetes' , 'cm_htn_c' , 'cm_hypothy' , 'cm_liver' , 'cm_obese' , 'cm_perivasc',
  'cm_renlfail' , 'cancer' , 'hosp_loc' , 'insu' , 'smoke' , 'year' ,
'dys' , 'priormi' , 'treatment' , 'priorpci' , 'car' ,
'pstroke', "ps", "linps")


bal.wts1 = NULL


bal.wts1 = dx.wts(x = df4$wts1, data = df4, vars = covlist, treat.var = "hhc",
                  estimand = 'ATT')
```




```{r}

# using cobalt package 

library(cobalt)

df4$att.weights = with(df4, hhc + (1-hhc)*ps/(1-ps))

a = bal.tab(hhc ~ age + aweekend + female + cm_chf + cm_chrnlung +  diabetes + cm_htn_c + cm_hypothy + cm_liver + cm_obese + cm_perivasc +
cm_renlfail + cancer + hosp_loc + insu + smoke + year +
dys + priormi + treatment + priorpci + car +
pstroke
  ,data = df4, weights = "att.weights",distance = "p.score", method = "weighting", un = TRUE)


```


```{r}
pre.szd = dfb$Diff.Un*100
post.szd = dfb$Diff.Adj*100

covnames = c('age' , 'aweekend' , 'female' , 'cm_chf' , 'cm_chrnlung' ,
  'diabetes' , 'cm_htn_c' , 'cm_hypothy' , 'cm_liver' , 'cm_obese' , 'cm_perivasc',
  'cm_renlfail' , 'cancer' , 'hosp_loc' , 'insu_medicaid', "insu_medicare" ,"insu_others","insu_private",  'smoke' , 'year' ,
'dys' , 'priormi' , 'treatment_cabg' ,"treatment_medical", "treatment_pci", 'priorpci' , 'car' ,
'pstroke')

# create loveplot

temp1 <- data.frame(pre.szd, post.szd, covnames)#combine the 3 variable into a data.base.
library(dplyr)
temp1<-temp1 %>% arrange(pre.szd)%>% mutate (covnames= factor(covnames,covnames)) #use dplyr mutate and arrange function to sort the covariates by pre.szd.
library(reshape2)
temp2<-melt(temp1,measure.vars=c("pre.szd","post.szd")) ##melt the database base to get a two level variable ("pre.szd","post.szd") that I will use in the plot.

p<-ggplot(temp2, aes(value, covnames, shape= variable,colour=variable)) ## the first step is tell ggplot that I want to use temp2 database to plot the (value) variable on x axis, covnames on y axis. Then I tell it to differentiate between the plotted geom based on (variable)  which is two level categorical variable ("pre.szd","post.szd").

p<-p +geom_point(size =3)+theme_bw()+labs(title="Standardized Difference Plot")+labs(x="Standardized Difference (%)")+labs(y="Covariates") # here, I tell ggplot that I want the values to be plotted as point (geom_point), I chose size 3, I change the background theme to black and white, then I label the axes.

p<-p+theme(legend.title=element_blank())+scale_colour_manual(values = c("pre.szd"="red", "post.szd"="green"),labels=c("Before Matching","After Matching"))+scale_shape_discrete(labels=c("Before Matching","After Matching")) ## here, I drop the legend title, I assign specific colors to the points, I rename the legend's labels.

p<-p + theme(legend.justification=c(1,0), legend.position=c(1,0)) + geom_vline(xintercept = 0,colour="blue") + geom_vline(xintercept = c(-10,10) ,colour="blue",linetype = "longdash") # I reposition the legend to right lower corner, I add the vertical lines.


p2 = p+ coord_cartesian(xlim = c(-60, 50)) + scale_x_continuous(breaks = seq(from = -200, to = 200, by = 10)) # Finally, I change x-axis limits and change the breaks to every 10.


```


```{r}


df4$discharge = ordered(df4$discharge, levels = c("home", "hhc"))

satt = survey::svydesign(data = df4, ids = ~hosp_nrd, weights = ~discwt + wts1, 
          strata = ~nrd_stratum, nest = T)

psatt = svyglm(readmit ~ hhc, design = satt, family = "binomial"(link = "logit"))

betaiptw = coef(psatt)

library(sandwich)

se = sqrt(diag(vcovHC(psatt, type = "HC")))

causalrd = (betaiptw[2])

lcl = (betaiptw[2] - 1.96*se[2])

ucl = (betaiptw[2] + 1.96*se[2])

adjust = c(lcl, causalrd, ucl)

adjust 
```

