---
title: "Untitled"
author: "Salil V Deo"
date: "November 22, 2019"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

# code for AMI and home health care and readmissions
# continue from the stata dataset developed 

library(haven);library(survey);library(srvyr);library(tidyverse);
library(MASS);library(arm);library(magrittr)

setwd("E:\\stemi_readmit\\")

# get data

df = read_stata("ami_readmit1.dta")

str(df)

keep = c("age", "aweekend", "died", "discwt", "dispuniform",  "dx1", "dx2", "dx3", "dx4", "dx5", "dx6", "dx7", "dx8", 
"dx9", "dx10",  "elective", "female", 
"hosp_nrd", "key_nrd", "los", 
"nrd_stratum",  "pay1", "pr1", "pr2", "pr3", 
"pr4", "pr5", "pr6", "pr7", "pr8", "pr9", "pr10",  "totchg", 
"year", "zipinc_qrtl",  
"cm_aids", "cm_alcohol", "cm_anemdef", "cm_arth", "cm_bldloss", 
"cm_chf", "cm_chrnlung", "cm_coag", "cm_depress", "cm_dm", "cm_dmcx", 
"cm_drug", "cm_htn_c", "cm_hypothy", "cm_liver", "cm_lymph", 
"cm_lytes", "cm_mets", "cm_neuro", "cm_obese", "cm_para", "cm_perivasc", 
"cm_psych", "cm_pulmcirc", "cm_renlfail", "cm_tumor", "cm_ulcer", 
"cm_valve", "cm_wghtloss", "hosp_bedsize", "hosp_urcat4", "hosp_ur_teach", 
 "indexevent", "daystoreadmission", "readmit", "cabg", "pci", 
"smoke", "dyslip", "priormi", "priorpci", "priorcabg", "carotid", 
"priorstroke", "prioricd")

df2 = df[,(keep)]

glimpse(df2)

df2$discharge[df2$dispuniform == 1]<- "home"
df2$discharge[df2$dispuniform == 5]<- "hhc"

thrombo <- as.character(c("9910")) # smoker

df2$thrombo <- apply(df2[paste0("pr", 1:10)], 1, 
                     function(x) any(x %in% thrombo)) %>% 
  as.numeric()

table(df2$thrombo)


keep2 = c("age", "aweekend", "died", "discwt", "dispuniform", 
"elective", "female", "hosp_nrd", "key_nrd", "los", "nrd_stratum", 
"pay1",  "zipinc_qrtl",  "cm_chf", "cm_chrnlung", 
 "cm_dm", "cm_dmcx",  "cm_htn_c", "cm_liver",
 "cm_renlfail",  "hosp_bedsize", "hosp_urcat4", "hosp_ur_teach", 
"indexevent", "daystoreadmission", "readmit", "cabg", "pci", 
"smoke", "dyslip", "priormi", "priorpci", "priorcabg", "carotid", 
"priorstroke", "prioricd", "discharge", "thrombo")

df3 = df2[,c(keep2)]

dmiss = map(df3, ~sum(is.na(.))) # determine missing data

# only zipinc_qrtl 1181 , pay1 91 elective 87 missing 

# create and clean some variables prior to analysis

table(df3$pay1)

df3$pay1 = as.factor(df3$pay1)

df3$insu[df3$pay1 == "1"]<- "medicare"
df3$insu[df3$pay1 == "2"]<- "medicaid"
df3$insu[df3$pay1 == "3"] <- "private"
df3$insu[df3$pay1 == "4"]<- "others"
df3$insu[df3$pay1 == "5"]<- "others"
df3$insu[df3$pay1 == "6"]<- "others"

table(df3$insu) # primary insurance 


# urban/rural

df3$hosp_loc[df3$hosp_ur_teach == "1"]<- "urban"
df3$hosp_loc[df3$hosp_ur_teach == "0"]<- "urban"
df3$hosp_loc[df3$hosp_ur_teach == "2"]<- "rural"


table(df3$hosp_loc) # urban/rural

table(df3$hosp_bedsize) # hospital bed-size
```



vars = c("age", "aweekend",  "elective", 
  "female",   
  "zipinc_qrtl", "cm_chf", "cm_chrnlung", "cm_dm", "cm_dmcx", "cm_htn_c", 
  "cm_liver", "cm_renlfail", "hosp_bedsize", 
  "indexevent",  "readmit", "cabg", "pci", 
  "smoke", "dyslip", "priormi", "priorpci", "priorcabg", "carotid", 
  "priorstroke", "prioricd", "discharge", "thrombo", "insu", "hosp_loc")

factorvars =  c( "aweekend",  "elective", 
"female",  "zipinc_qrtl", "cm_chf", "cm_chrnlung", "cm_dm", "cm_dmcx", "cm_htn_c", 
"cm_liver", "cm_renlfail", "hosp_bedsize", 
 "indexevent", "readmit", "cabg", "pci", 
 "smoke", "dyslip", "priormi", "priorpci", "priorcabg", "carotid", 
"priorstroke", "prioricd", "discharge", "thrombo", "insu", "hosp_loc")

```{r}
options(survey.lonely.psu = "adjust")

amis = survey::svydesign(data = df3, strata = ~nrd_stratum,
id = ~hosp_nrd, weight = ~discwt, nest = T)
```


```{r}
# for cabg

cabgmi = subset(amis, cabg == 1)

cabg = prop.table(svytable(~readmit + discharge, design = cabgmi), 2)

svychisq(~readmit + discharge, design = cabgmi)

# for pci

pcimi = subset(amis, pci == 1)

pci = prop.table(svytable(~readmit + discharge, design = pcimi), 2)

svychisq(~readmit + discharge, design = pcimi)

pci


# bar plot

barp = read_csv("barp.csv")

library(ggthemes)

p = ggplot(data = barp, aes(x = Procedure, y = Percentage, fill = Category), alpha = 0.5) + geom_bar(stat = "identity", position = "dodge", size = 10) + scale_fill_grey(start = 0.4, end = 0.8)

p2 = p + ggthemes::theme_clean() + xlab("Procedure Performed") + ylab("Percentage of Patients Readmitted")

```


# t1 = tableone::svyCreateTableOne(data = amis, vars = vars,
# factorVars = factorvars, strata = "discharge") # create tableone for entire
# data

# table1 = print(t1)

# write.csv(table1, "table1.csv")

# plan to do unadjusted regression for readmission with home health care.
# then plan for propensity matching and then reanalysis.

```{r}
# unadjusted analysis

dfs = survey::svydesign(data = df3, 
strata = ~nrd_stratum, 
weight = ~discwt, ids  = ~hosp_nrd, nest = T)

options(survey.lonely.psu = "adjust")

# unadjusted survey regression for readmission

readmit_r = svyglm(readmit ~ discharge, design = dfs)

summary(readmit_r)

exp(coef(readmit_r))
exp(confint(readmit_r))


### peforming propensity matched analysis for readmission after home health care


library(ggplot2); library(dplyr); library(Matching)
library(Hmisc);  library(Epi); library(lme4)
library(survival); library(twang); library(tableone)
library(survey);library(tidyverse);library(magrittr)
library(purrr);library(haven)


library(purrr)

miss = purrr::map(df3, ~sum(is.na(.)))

# plan to simple imputation with adding most common category for missing data
# insu, zipinc_qrtl, elective

table(df3$insu, useNA = "ifany")

table(df3$zipinc_qrtl, useNA = 'ifany')

table(df3$elective, useNA = 'ifany')

df3$insu[is.na(df3$insu)]<- "medicare"
df3$elective[is.na(df3$elective)]<- 0
df3$zipinc_qrtl[is.na(df3$zipinc_qrtl)]<- 2

# no missing data now


keep3 = c("age", "aweekend",  "discwt",  "elective", 
  "female", "hosp_nrd", "key_nrd", "nrd_stratum",  
  "zipinc_qrtl", "cm_chf", "cm_chrnlung", "cm_dm", "cm_dmcx", "cm_htn_c", 
  "cm_liver", "cm_renlfail", "hosp_bedsize",  
   "readmit", "cabg", "pci", 
  "smoke", "dyslip", "priormi", "priorpci", "priorcabg", "carotid", 
  "priorstroke", "prioricd", "discharge", "thrombo", "insu", "hosp_loc")

propdata = df3[,c(keep3)]

propdata$disc_n[propdata$discharge == "home"]<- 0
propdata$disc_n[propdata$discharge == "hhc"]<- 1

props = survey::svydesign(data = propdata, 
strata = ~nrd_stratum, 
weight = ~discwt, ids  = ~hosp_nrd, nest = T)

# create logistic regression model for propensity for use of home health care

psmodel = svyglm(disc_n ~ age + factor(aweekend) + factor(female) + 
factor(zipinc_qrtl) + factor(cm_chf) + factor(cm_chrnlung) + factor(cm_dm) +
factor(cm_dmcx) + factor(cm_renlfail) + factor(cm_liver) + factor(cm_htn_c) + factor(hosp_bedsize) +
factor(cabg) + factor(pci) + factor(priormi) + factor(priorpci) + 
factor(priorcabg) + factor(priorstroke) + factor(prioricd) + factor(thrombo) +
insu + hosp_loc, design = props, family = "binomial")

summary(psmodel)


propdata$ps <- psmodel$fitted
propdata$linps <- psmodel$linear.predictors

library(ggthemes)

p = ggplot(propdata, aes(x = ps, fill = discharge)) +
  geom_density(alpha=0.3) +
  labs(x = "Propensity Score", colour="Type", shape="Type",
       y = "Distribution Density") + ylim(0,5) + 
  ggthemes::theme_clean() + 
  scale_fill_manual(values = c("blue","red"))


X <- propdata$linps
Tr <- as.logical(propdata$disc_n)


match <- Match(Tr=Tr, X=X, M = 1, replace=FALSE, ties=FALSE,
               caliper = 0.5)
summary(match)


set.seed(50001) # set seed for bootstrap assesssments - not critical
mb1 <- MatchBalance(disc_n ~ age + factor(aweekend) + factor(female) + 
                      factor(zipinc_qrtl) + factor(cm_chf) + factor(cm_chrnlung) + factor(cm_dm) +
                      factor(cm_dmcx) + factor(cm_renlfail) + factor(cm_liver) + factor(cm_htn_c) + factor(hosp_bedsize) +
                      factor(cabg) + factor(pci) + factor(priormi) + factor(priorpci) + 
                      factor(priorcabg) + factor(priorstroke) + factor(prioricd) + factor(thrombo) +
                      insu + hosp_loc,data = propdata,
                    match.out = match, nboots=500)


covnames <- c("Age", "Weekend admit", "Female", "Median income",
"CHF", "COPD", "DM", "Comp_DM", "CKD", "Liver disease","HTN", "Bedsize",
"CABG", "PCI", "Prior MI", "Prior PCI", "Prior CABG", "Prior Stroke", "Prior ICD", "Thrombolysis",
"Insurance", "Hospital Location")



pre.szd <- NULL; post.szd <- NULL
for(i in 1:length(covnames)) {
  pre.szd[i] <- mb1$BeforeMatching[[i]]$sdiff.pooled
  post.szd[i] <- mb1$AfterMatching[[i]]$sdiff.pooled
}

temp <- data.frame(pre.szd, post.szd, row.names=covnames)
print(round(temp,1))

              

temp <- data.frame(pre.szd, post.szd, row.names=covnames)
tempsort <- temp[with(temp, order(abs(pre.szd))),]
high <- max(max(abs(pre.szd)), max(abs(post.szd)), 0.1)

dotchart(abs(tempsort$pre.szd), pch="", 
         xlim=c(0, 1.05*high), labels=row.names(tempsort), cex= 1,fontsize = 5,
         main="Absolute Standardized Difference", 
         xlab="Absolute Standardized Difference (%)")
points(abs(tempsort$pre.szd), seq(1:length(tempsort$pre.szd)), 
       pch=15, col="gray", cex=1)
points(abs(tempsort$post.szd), seq(1:length(tempsort$post.szd)), 
       pch=19, col="red", cex=1)
abline(v=0, lty=1)
abline(v=10, lty=2, col="purple")
legend("bottomright", legend = c("Before Match", "After Match"), 
       col=c("gray", "red"), text.col=c("gray", "red"), 
       bty="o", pch = c(15, 19))

```


```{r}
temp <- data.frame(pre.szd, post.szd, row.names=covnames)


temp1 <- data.frame(pre.szd, post.szd, covnames)#combine the 3 variable into a data.base.
library(dplyr)
temp1<-temp1 %>% arrange(pre.szd)%>% mutate (covnames= factor(covnames,covnames)) #use dplyr mutate and arrange function to sort the covariates by pre.szd.
library(reshape2)
temp2<-melt(temp1,measure.vars=c("pre.szd","post.szd")) ##melt the database base to get a two level variable ("pre.szd","post.szd") that I will use in the plot.

p<-ggplot(temp2, aes(value, covnames, shape= variable,colour=variable)) ## the first step is tell ggplot that I want to use temp2 database to plot the (value) variable on x axis, covnames on y axis. Then I tell it to differentiate between the plotted geom based on (variable)  which is two level categorical variable ("pre.szd","post.szd").

p<-p +geom_point(size =3)+theme_bw()+labs(title="Standardized Difference Plot")+labs(x="Standardized Difference (%)")+labs(y="Covariates") # here, I tell ggplot that I want the values to be plotted as point (geom_point), I chose size 3, I change the background theme to black and white, then I label the axes.

p<-p+theme(legend.title=element_blank())+scale_colour_manual(values = c("pre.szd"="red", "post.szd"="blue"),labels=c("Before Matching","After Matching"))+scale_shape_discrete(labels=c("Before Matching","After Matching")) ## here, I drop the legend title, I assign specific colors to the points, I rename the legend's labels.

p<-p + theme(legend.justification=c(1,0), legend.position=c(1,0)) + geom_vline(xintercept = 0,colour="blue") + geom_vline(xintercept = c(-10,10) ,colour="blue",linetype = "longdash") # I reposition the legend to right lower corner, I add the vertical lines.


p2 = p+ coord_cartesian(xlim = c(-60, 50)) + scale_x_continuous(breaks = seq(from = -200, to = 200, by = 10)) # Finally, I change x-axis limits and change the breaks to every 10.
```



```{r}


matches <- factor(rep(match$index.treated, 2))

data.m1sample <- cbind(matches, 
propdata[c(match$index.control, match$index.treated),])


data.m1sample = tbl_df(data.m1sample) # create tibble
```
### write_csv(data.m1sample, "propdata.csv")


vars = c("age", "aweekend",  "elective", 
  "female",   
  "zipinc_qrtl", "cm_chf", "cm_chrnlung", "cm_dm", "cm_dmcx", "cm_htn_c", 
  "cm_liver", "cm_renlfail", "hosp_bedsize", 
  "indexevent",  "readmit", "cabg", "pci", 
  "smoke", "dyslip", "priormi", "priorpci", "priorcabg", "carotid", 
  "priorstroke", "prioricd", "discharge", "thrombo", "insu", "hosp_loc")

factorvars =  c( "aweekend",  "elective", 
"female",  "zipinc_qrtl", "cm_chf", "cm_chrnlung", "cm_dm", "cm_dmcx", "cm_htn_c", 
"cm_liver", "cm_renlfail", "hosp_bedsize", 
 "indexevent", "readmit", "cabg", "pci", 
 "smoke", "dyslip", "priormi", "priorpci", "priorcabg", "carotid", 
"priorstroke", "prioricd", "discharge", "thrombo", "insu", "hosp_loc")


```{r}

options(survey.lonely.psu = "adjust")

props = survey::svydesign(data = data.m1sample, strata = ~nrd_stratum,
id = ~hosp_nrd, weight = ~discwt, nest = T)
```
t2 = tableone::svyCreateTableOne(data = props, vars = vars,
factorVars = factorvars, strata = "discharge") # create tableone for entire
# data

table2 = print(t2)

write.csv(table2, "table2.csv")


```{r}


#### 
# including all variables tha are different 

R = svyglm(readmit ~ factor(disc_n) + priorcabg + smoke + dyslip + elective + age, design = props,
           family = "binomial")

summary(R)

exp(coef(R))
exp(confint(R))


             
#### the model above is the one written and used in the paper 
# it adjusts further for all the variables that were different in table 2.




```


```{r}

# descriptive statistics of the whole cohort 
# df3 is the data

options(survey.lonely.psu = "adjust")

desc = survey::svydesign(data = df3, strata = ~nrd_stratum,
id = ~hosp_nrd, weight = ~discwt, nest = T)

svymean(~age, desc)

svymean(~female, desc)

svymean(~cm_renlfail, desc)

svymean(~cm_dm, desc)

svymean(~cm_dmcx, desc)

svymean(~priormi, desc)

svymean(~priorcabg, desc)

svymean(~priorpci, desc)

svymean(~cabg, desc)

svymean(~pci, desc)

svymean(~thrombo, desc)

svymean(~dyslip, desc)

svymean(~cm_htn_c, desc)

table(df3$hosp_loc)

prop.table((svytable(~hosp_loc, desc)))

prop.table(svytable(~discharge, desc))

svymean(~readmit, desc)
```


```{r}


# discuss Dr Kapadia`s question: Does CABG confound findings here ?
# CABG has been adjusted for in the propensity model, so the final result considers CABG also
# CABG only cohort, HHC and readmission. ? adjust for other factors ?
# PCI only cohort, HHC and readmission. - ? adjust for other factors ?

# initially create a seperate table for only patients with PCI done

dfpci = data.m1sample %>% filter(pci == 1)
str(dfpci)

# create tableone for this data

vars = c("age", "aweekend",  "elective", 
  "female",   
  "zipinc_qrtl", "cm_chf", "cm_chrnlung", "cm_dm", "cm_dmcx", "cm_htn_c", 
  "cm_liver", "cm_renlfail", "hosp_bedsize", 
  "indexevent",  "readmit", "cabg", "pci", 
  "smoke", "dyslip", "priormi", "priorpci", "priorcabg", "carotid", 
  "priorstroke", "prioricd", "discharge", "thrombo", "insu", "hosp_loc")

factorvars =  c( "aweekend",  "elective", 
"female",  "zipinc_qrtl", "cm_chf", "cm_chrnlung", "cm_dm", "cm_dmcx", "cm_htn_c", 
"cm_liver", "cm_renlfail", "hosp_bedsize", 
 "indexevent", "readmit", "cabg", "pci", 
 "smoke", "dyslip", "priormi", "priorpci", "priorcabg", "carotid", 
"priorstroke", "prioricd", "discharge", "thrombo", "insu", "hosp_loc")


options(survey.lonely.psu = "adjust")

pci_survey = survey::svydesign(data = dfpci, strata = ~nrd_stratum,
id = ~hosp_nrd, weight = ~discwt, nest = T)

table_pci = tableone::svyCreateTableOne(data = pci_survey , vars = vars,
factorVars = factorvars, strata = "discharge") # create tableone for entire
# data

table2 = print(table_pci)

# from this table the differences are:
# elective, cm_chf, cm_renlfail, hosp_bedsize, smoke, dyslip, priorcabg


pcionly = svyglm(readmit ~ discharge + priorcabg + smoke + dyslip + elective + cm_chf +
                   cm_renlfail + hosp_bedsize, design = pci_survey,
           family = "binomial")

summary(pcionly)

exp(coef(pcionly))
exp(confint(pcionly))



```




